##Efficacy analytics audition project
Purpose of this project is to explore the related dataset, and prepare a
presentation of any findings and/or actionable insights.

List of assumptions I took during the analysis:

* The missing values are missing at random, that is - value of the variable or any other variable is not related to the fact if the value is input properly or not

* When analyzing mean of average score for different countries, I excluded the countries for which number of student is lower than or equal to 50. That is because it may not be right to compare average for 5 students in one group with average of 1000 students in the other.

* The same student can have two units assigned at the same time (learning two chapters at once), but cannot have two different country codes assigned 

The whole process can be divided into sections:  
1. Uploading data and data quality check  
2. Assessing colinearity  
3. Trying to fit linear model to data  
4. Calculating mean of student scores for different countries  
5. Analysis of mean for groups of students, grouped by country and variable in_course

Conclusions will be presented in the end of this file.

1. Uploading data and data quality checks
```{r eval=TRUE}
EAdata <- read.csv("D:/Nauka/R/Pearson zadanie/Audition project/EA_audition_for_sharing/data.csv")
```
Quick look at the data summary and variables classes
```{r eval=TRUE}
summary(EAdata)
nrow(EAdata)
```
```{r eval=TRUE}
for (x in 1:length(EAdata)) {
  print( class(EAdata[,x])  )
}
```
I performed two simple checks, to know if:

+ There are any duplicated rows in the data

+ There are any rows where learner with particular id has more than one country assigned
```{r eval=TRUE}
dup<-duplicated(EAdata)
summary(dup)

ids<-unique(EAdata[,1]) 
for (x in ids) {
  tmp<-unique(EAdata[EAdata$learner_id==x,2])
  if (length(tmp)!=1) {
    print(EAdata[EAdata$learner_id==x,])
  }
}
```
Both checks did not show any issues. The dataset summary also did not depict any issues, so we can move on with the analysis.
  
2. Checking correlations for different variables
```{r}
EAdata.3.5<-na.omit(EAdata[,c(3,5)])
cor(as.numeric(EAdata.3.5[,1]),EAdata.3.5[,2])

EAdata.5.7<-na.omit(EAdata[,c(5,7)])
cor(as.numeric(EAdata.5.7[,1]),EAdata.5.7[,2])

EAdata.4.5<-na.omit(EAdata[,c(4,5)])
cor(as.numeric(EAdata.4.5[,1]),EAdata.4.5[,2])
```
Looks like there is no statisticaly relevant colinearity between avg_score and the variables - inv_rate,unit & in_course

  3. Trying to fit linear model to data
```{r results='hide'}
model1 <- lm(formula = avg_score ~ as.numeric(in_course) + inv_rate + as.numeric(country), 
            data = EAdata)
modelStep <- step(lm(formula = avg_score ~ ., data = EAdata[,-1]))
model2 <- lm(formula = avg_score ~ in_course,  data = EAdata)
```
```{r echo=FALSE}
summary(model2)
```
None of the models looked useful. Although t-test showed some variables as relevant, there was R^2 statistics showed very weak fit

  4. Calculating mean of student scores for different countries
```{r results='hide', message=FALSE, warning=FALSE}
##Calculating mean for of avg_score for each country
library(dplyr)
meanScores<- EAdata %>%
  group_by(country) %>%
  summarise_each(funs(mean(., na.rm=TRUE)), avg_score)
##Let's add to this data frame column with number of student
studentsCountries <- as.data.frame( table(EAdata$country) )
meanScores <- cbind(meanScores, studentsNumber=studentsCountries$Freq)

##Table of mean scores in each country (only those with more than 50 students)
df_meanScores <- as.data.frame(meanScores[meanScores$studentsNumber>50,])
df_meanScores = df_meanScores[ order(df_meanScores$avg_score), ]
```
It turns out, that after filtering out countries number of students higher than 50, we get only `r length(df_meanScores$country)` countries. Without filtering, it was `r length(meanScores$country)`.

Once we have the table of mean scores in filtered countries, we can have a look at it's top and bottom
```{r}
head(df_meanScores)
tail(df_meanScores)
```
We may conclude, that in this ranking students from Ukraine showed the best average score, and students from Great Britian the worst average score

Presenting the top five countries on a bar plot:
```{r warning=FALSE}
selCountries1000 <- meanScores[meanScores$studentsNumber>1000,1]
EAdataFiltered <- EAdata[EAdata$country == selCountries1000,]
plt<-barplot(table(EAdataFiltered$in_course,as.character( EAdataFiltered$country) ),beside = TRUE, col=c("red","yellow"))
legend("topleft", c("studying alone","studying in course"), lwd=10, col=c("red","yellow"))
tab<-table(EAdataFiltered$in_course,as.character( EAdataFiltered$country) )
```
There are only five countries with number of students higher than 1000.
Five countries with the highest number of students (total) are presented on the plot.

  5. Analysis of mean for groups of students, grouped by country and variable in_course
```{r}  
testMean <- aov(avg_score~country, data = EAdataFiltered)
summary(testMean)
TukeyHSD(testMean) #Comparing mean score for each of 5 countries with highest number of students.

```
All the tested pairs of countries has significantly different mean score from each other, except of IT-ES (for 95% confidence level).
It is safe to conclude that average scores are different between countries.
Are these different for students studying alone and studying with teacher?
```{r warning=FALSE}
selCountries50 <- meanScores[meanScores$studentsNumber>50,1]
EAdataFilteredSmall <- EAdata[EAdata$country == selCountries50,]
testMean <- aov(avg_score~in_course, data = EAdataFilteredSmall)
summary(testMean)
summary(EAdataFilteredSmall[EAdataFilteredSmall$in_course=="t",5])
summary(EAdataFilteredSmall[EAdataFilteredSmall$in_course=="f",5])
```
Again, Anova results show that the difference between groups studying alone and studying in course is statistically significant (for 95% confidence level).

If we calculate mean of average scores for all the students, how many students will have a higher avg_score than than the mean for all students?
(I am not filtering student numbers here, because I will be not grouping by countries this time)
```{r}
totalMean<-mean(EAdata$avg_score, na.rm = TRUE)
compMeanData<-EAdata[EAdata$avg_score>totalMean,]
print(totalMean)
median(EAdata$avg_score, na.rm = TRUE)
```
Out of `r length(unique( EAdata$learner_id ))` there were `r length(unique( compMeanData$learner_id ))` students with the average score higher than the mean for all students. Besides, median of avg_score is higher than mean of avg_score, so we can expect the distribution to be negatively skewed.

### Conclusions:

* **Avg_score is not dependent on any of these variables - inv_rate,unit & in_course showed. This is consistent with the results of fitting any linear model. **Probably there are other important factors telling if the student will have a high score or not.
* **There were students out of 87 countries in total, but there are 57 countries in the data where number of students is less than or equal to 50 students.** Maybe these countries should be checked in terms of profitability?
* Looking at the countries with more than 50 students provided - **there is a significant difference between the average score for each whole country. The country with the lowest average score turned out to be Great Britian, the highest - Ukraine.** It would be worth considering preparing more profiled courses for countries with the lowest scores.
* **Average score for group of students learning with teacher turned out to be significantly higher than the average score for group of students learning on their own.** If students are to receive higher scores, more profiled elearning courses should be considered.
* The distribution of avg_score values is negatively skewed, which means that there are more students receiving high scores than those receiving lower scores. **About `r round(length(unique( compMeanData$learner_id ))/length(unique( EAdata$learner_id )) ,2)` % of students had the average score higher than the mean for all students.**
